---
title: "4.1 GAM Predictive Models"
author: "Colby Cooper"
date: "`r Sys.Date()`"
output: html_document
---

```{r document_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages and data and setup base variables

```{r data_setup, warnings=FALSE, message=FALSE}
# Packages
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(mgcv)

# Load data
journeys <- read.csv("customer_journey-4.csv")
results <- read.csv("results-3.csv")

# Channels to keep order consistent
channels <- c("Direct",
              "Email Marketing",
              "Facebook",
              "Google Display",
              "Google Search",
              "Instagram",
              "Organic",
              "Youtube")

# States for predictive models 1 and 2
states_1 <- c(channels,
              "Conversion",
              "Exit")

states_2 <- c(channels,
              "Conversion")

# Number of basis functions for the GAM smooth
gam_k <- 6

# Maximum step to which we will generate predictions
max_prediction_step <- 30

# Set seed for reproducibility
set.seed(123)
```

Combine datasets into one table

```{r combine_datasets}
journeys_full <- journeys %>%
  left_join(results %>%
              select(user_id, session_id, conversion),
            by = c("user_id", "session_id")) %>%
  arrange(user_id, session_id, order)
```

Split data into training and test groups

```{r training_test_split}
# Get all unique customers
all_users <- unique(journeys_full$user_id)

# Randomly select 80% of customers for training
training_users <- sample(all_users,
                  size = floor(0.80 * length(all_users)))

# Remaining 20% of customers for testing
test_users <- setdiff(all_users, training_users)

# Split journey data
journeys_training <- journeys_full %>%
  filter(user_id %in% training_users)

journeys_test <- journeys_full %>%
  filter(user_id %in% test_users)
```

Create function to count transitions and use for both models

```{r transitions}
# Function to extract transitions data
make_transitions <- function(data,
                             return_only = FALSE,
                             converting_only = FALSE) {
  
  x <- data
  
  # Model 1: only return visits
  if (return_only) {
    x <- x %>%
      filter(session_id > 0)
  }
  
  # Model 2: only converting journeys
  if (converting_only) {
    x <- x %>%
      filter(conversion == 1)
  }
  
  x %>%
    arrange(user_id, session_id, order) %>%
    group_by(user_id, session_id) %>%
    mutate(
      
      current_state = channel,
      
      next_state = lead(channel),
      
      # Final touchpoint of each session:
      # if conversion = 1 -> Conversion
      # otherwise -> Exit
      next_state = case_when(
        !is.na(next_state) ~ next_state,
        conversion == 1 ~ "Conversion",
        TRUE ~ "Exit"),
      
      # Journey stage
      step = order) %>%
    ungroup() %>%
    select(user_id,
           session_id,
           step,
           current_state,
           next_state,
           conversion)
}

# Model 1 transitions
transitions_1 <- make_transitions(journeys_training,
                                   return_only = TRUE)

# Model 2 transitions
transitions_2 <- make_transitions(journeys_training,
                                   converting_only = TRUE)
```

Aggregate transitions to make calculations faster

```{r aggregate}
# Model 1
transition_counts_1 <- transitions_1 %>%
  count(step,
        current_state,
        next_state,
        name = "Count")

# Model 2
transition_counts_2 <- transitions_2 %>%
  count(step,
        current_state,
        next_state,
        name = "Count")
```

Fit GAMs to every possible transition

```{r GAMs}
# Function to fit GAMs to transitions
fit_transition_gams <- function(data,
                                states,
                                channels,
                                k = k) {
  
  curves <- list()
  
  for (from_state in channels) {
    
    curves[[from_state]] <- list()
    
    # Total number of transitions originating from this state at each step
    
    totals <- data %>%
      filter(.data$current_state == from_state) %>%
      group_by(step) %>%
      summarise(
        total = sum(Count),
        .groups = "drop")
    
    # Fit one model for every possible destination

    for (to_state in states) {
      
      successes <- data %>%
        filter(.data$current_state == from_state,
               .data$next_state == to_state) %>%
        group_by(step) %>%
        summarise(success = sum(Count),
                  .groups = "drop")
      
      # Add zero successes for steps where this transition did not occur
      d <- totals %>%
        left_join(successes,
                  by = "step") %>%
        mutate(success = replace_na(success, 0),
               failure = total - success)
      
      # Special cases

      # Transition never occurs
      if (sum(d$success) == 0) {
        
        curves[[from_state]][[to_state]] <- list(
          type = "constant",
          value = 0)
        
        # Transition always occurs
      } else if (sum(d$success) == sum(d$total)) {
        
        curves[[from_state]][[to_state]] <- list(
          type = "constant",
          value = 1)
        
        # Normal GAM

      } else {
        
        fit <- gam(
          cbind(success, failure) ~
            s(step, k = k),
          family = binomial,
          method = "REML",
          data = d)
        
        curves[[from_state]][[to_state]] <- list(
          type = "gam",
          fit = fit)
      }
    }
  }
  
  list(curves = curves,
       states = states,
       channels = channels,
       k = k)
}

# Produce GAMs for model 1
model_1 <- fit_transition_gams(
  data = transition_counts_1,
  states = states_1,
  channels = channels,
  k = gam_k)

# Produce GAMs for model 2
model_2 <- fit_transition_gams(
  data = transition_counts_2,
  states = states_2,
  channels = channels,
  k = gam_k)
```

Function to use transition curve to predict transition probability

```{r predict_transition}
predict_transition_curve <- function(curve,
                                     new_step) {
  
  # Constant model
  if (curve$type == "constant") {
    
    return(rep(curve$value,
               length(new_step)))
  }
  
  # GAM
  p <- predict(curve$fit,
               newdata = data.frame(step = new_step),
               type = "response")
  
  # Protect against tiny numerical values outside [0,1]
  pmin(pmax(p, 0),1)
}
```

Extract GAM curves

```{r extract_curves}
# Function to extract curves
extract_curves <- function(model,
                           prediction_steps) {
  
  output <- list()
  
  counter <- 1
  
  for (from_state in model$channels) {
    
    for (to_state in model$states) {
      
      curve <- model$curves[[from_state]][[to_state]]
      
      probabilities <- predict_transition_curve(
        curve,
        prediction_steps)
      
      output[[counter]] <- data.frame(
        from_state = from_state,
        to_state = to_state,
        step = prediction_steps,
        probability = probabilities)
      
      counter <- counter + 1
    }
  }
  
  bind_rows(output)
}

# Set steps for prediction
prediction_steps <- 0:max_prediction_step

# Model 1 GAM curves
curve_predictions_1 <- extract_curves(
  model_1,
  prediction_steps)

# Model 2 GAM curves
curve_predictions_2 <- extract_curves(
  model_2,
  prediction_steps)
```

Example GAMs for each model (not transition probability)

```{r GAM_plots}
# Model 1 plot
example_channel_1 <- "Instagram"

example_plot_1 <- curve_predictions_1 %>%
  filter(from_state == example_channel_1) %>%
  ggplot(aes(x = step,
             y = probability,
             colour = to_state)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = max(transitions_1$step),
             linetype = "dashed") +
  labs(title = paste("Model 1 Smoothed Transition Probabilities From",
                     example_channel_1),
       subtitle = "Dashed Line Indicates the Maximum Observed Journey step",
       x = "Current Step",
       y = "GAM Predicted Transition Probability",
       colour = "Next state") +
  theme_minimal()

print(example_plot_1)

# Model 2 plot
example_channel_2 <- "Instagram"

example_plot_2 <- curve_predictions_2 %>%
  filter(from_state == example_channel_2) %>%
  ggplot(aes(x = step,
             y = probability,
             colour = to_state)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = max(transitions_2$step),
             linetype = "dashed") +
  labs(title = paste("Model 2 Smoothed Transition Probabilities From",
                     example_channel_2),
       subtitle = "Dashed Line Indicates the Maximum Observed Journey step",
       x = "Current Step",
       y = "GAM Predicted Transition Probability",
       colour = "Next State") +
  theme_minimal()

print(example_plot_2)
```

Function to create transition probability matrix
```{r make_matrix}
make_matrix <- function(model,
                        step) {
  
  states <- model$states
  
  M <- matrix(0,
              nrow = length(states),
              ncol = length(states),
              dimnames = list(states,
                              states))
  
  # Non-absorbing states

  for (from_state in model$channels) {
    
    probabilities <- sapply(
      states,
      function(to_state) {
        
        curve <- model$curves[[from_state]][[to_state]]
        
        predict_transition_curve(
          curve,
          new_step = step
        )[1]
      }
    )
    
    probabilities <- as.numeric(probabilities)
    
    # Normalise row

    probability_sum <- sum(probabilities)
    
    if (probability_sum > 0) {
      
      probabilities <-
        probabilities / probability_sum
      
    } else {
      
      stop(paste("All transition probabilities are zero for",
                 from_state,
                 "at step",
                 step))
    }
    
    M[from_state, ] <- probabilities
  }
  
  # Absorbing states

  absorbing_states <- setdiff(states,
                              model$channels)
  
  for (state in absorbing_states) {
    
    M[state, ] <- 0
    M[state, state] <- 1
  }
  
  
  M
}
```

Example plots of transition probabilities (normalised GAM values)

```{r transition_probabilities_plots}
# Extract normalised curves for plotting
extract_normalised_curves <- function(model, prediction_steps) {
  
  output <- list()
  counter <- 1
  
  for (step in prediction_steps) {
    
    M <- make_matrix(model, step)
    
    for (from_state in model$channels) {
      
      for (to_state in model$states) {
        
        output[[counter]] <- data.frame(
          from_state = from_state,
          to_state = to_state,
          step = step,
          probability = M[from_state, to_state]
        )
        
        counter <- counter + 1
      }
    }
  }
  
  bind_rows(output)
}

# Model 1 transition probabilities
max_observed_step_1 <- max(transitions_1$step, na.rm = TRUE)

normalised_curves_1 <- extract_normalised_curves(
  model_1,
  prediction_steps)

curves_1 <- normalised_curves_1 %>%
  filter(from_state == example_channel_1)

ggplot(curves_1,
       aes(x = step, y = probability, colour = to_state)) +
       geom_line(linewidth = 1) +
       geom_vline(xintercept = max_observed_step_1,
             linetype = "dashed") +
  labs(title = paste("Model 1 Transition Probabilities From",
                     example_channel_1),
       subtitle = "Dashed Line Indicates the Maximum Observed Journey step",
       x = "Current Step",
       y = "Transition Probability",
       colour = "Next State") +
  theme_minimal()

# Model 2 transition probabilities
max_observed_step_2 <- max(transitions_2$step, na.rm = TRUE)

normalised_curves_2 <- extract_normalised_curves(
  model_2,
  prediction_steps)

curves_2 <- normalised_curves_2 %>%
  filter(from_state == example_channel_2)

ggplot(curves_2,
       aes(x = step, y = probability, colour = to_state)) +
       geom_line(linewidth = 1) +
       geom_vline(xintercept = max_observed_step_2,
             linetype = "dashed") +
  labs(title = paste("Model 2 Transition Probabilities From",
                     example_channel_2),
       subtitle = "Dashed Line Indicates the Maximum Observed Journey step",
       x = "Current Step",
       y = "Transition Probability",
       colour = "Next State") +
  theme_minimal()
```

Build the necessary transition matrices based on data and steps

```{r matrices}
# Function to build necessary transition matrices
build_matrices <- function(model,
                           max_step) {
  
  matrices <- lapply(
    0:max_step,
    function(n) {
      make_matrix(
        model,
        step = n)
    }
  )
  
  names(matrices) <- as.character(0:max_step)
  
  matrices
}

# Build matrices for model 1
matrices_1 <- build_matrices(model_1,
                             max_prediction_step)

# Build matrices for model 2
matrices_2 <- build_matrices(model_2,
                             max_prediction_step)
```

# Model 1 - Conversion Probability On Return Visit

Set up initial state probability vector

```{r initials_1}
# Initial state counts and probabilities
return_start <- transitions_1 %>%
  filter(step == 0) %>%
  count(current_state,
        name = "count") %>%
  mutate(probability = count / sum(count))

# Set initial states vector
p0_1 <- setNames(rep(0,length(states_1)),
                 states_1)

p0_1[return_start$current_state] <- return_start$probability
```

Propagate initial states through the Markov chain

```{r propagate_1}
p <- p0_1

distribution_1 <- list()

for (step in 0:max_prediction_step) {
  
  distribution_1[[step + 1]] <- data.frame(step = step,
                                           state = names(p),
                                           probability = as.numeric(p))
  
  # Stop if this is the final requested step
  if (step == max_prediction_step) {
    break
  }
  
  M <- matrices_1[[as.character(step)]]
  
  p <- as.numeric(p %*% M)
  
  names(p) <- states_1
}

distribution_1 <- bind_rows(distribution_1)
```

Calculate cumulative conversion probability

```{r cumulative_conversion}
conversion_probability_1 <- distribution_1 %>%
  filter(state == "Conversion") %>%
  select(step,
         conversion_probability = probability)

# View the first few cumulative conversion probabilities
print(head(conversion_probability_1))
```

Plot the cumulative conversion probability of return journeys

```{r cumulative_conversion_plot}
conversion_plot_1 <- ggplot(conversion_probability_1,
                            aes(x = step,
                                y = conversion_probability))+ 
  geom_line(linewidth = 1) +
  geom_vline(xintercept = max(transitions_1$step),
             linetype = "dashed") +
  scale_y_continuous(limits = c(0, 1),
                     labels = scales::percent) +
  labs(title = "Model 1 Predicted Cumulative Probability of Conversion",
       subtitle = "Return-Visit Journeys",
       x = "Journey Step",
       y = "Cumulative Probability of Conversion") +
  theme_minimal()

print(conversion_plot_1)
```

Calculate how conversion probability changes depending on starting channel

```{r conversion_by_channel}
# Function to calculate conversion probability by starting channel
conversion_probability_from_state <- function(
    model,
    matrices,
    start_state,
    start_step = 0,
    end_step = max_prediction_step) {
  
  p <- setNames(
    rep(0,
        length(model$states)),
    model$states)
  
  p[start_state] <- 1
  
  for (step in start_step:end_step) {
    
    if (step == end_step) {
      break
    }
    
    M <- matrices[[as.character(step)]]
    
    p <- as.numeric(p %*% M)
    
    names(p) <- model$states
  }
  
  p["Conversion"]
}

# Conversion rate by initial channel
channel_conversion_1 <- tibble(
  channel = channels,
  conversion_probability = map_dbl(
    channels,
    ~ conversion_probability_from_state(
      model_1,
      matrices_1,
      start_state = .x,
      start_step = 0,
      end_step = max_prediction_step)))

print(channel_conversion_1)
```

# Model 2 - Expected Steps Until Conversion

E[n | current channel, current step, eventual conversion]

Function to calculate expected steps until conversion

```{r remaining_steps_function}
expected_steps_to_conversion <- function(
    model,
    matrices,
    start_state,
    start_step = 0,
    end_step = max_prediction_step) {
  
  # Initial state

  p <- setNames(rep(0,
                    length(model$states)),
                model$states)
  
  p[start_state] <- 1
  
  total_conversion_probability <- 0
  
  weighted_steps <- 0
  
  # Propagate forward

  if (start_step >= end_step) {
    return(NA_real_)
  }
  
  for (step in start_step:end_step) {
    
    M <- matrices[[as.character(step)]]
    
    p_new <- as.numeric(p %*% M)
    
    names(p_new) <- model$states
    
    # Probability that conversion occurs for the first time on this transition. Since Conversion is absorbing, this is simply the increase in the probability of being in Conversion
    
    conversion_this_step <-
      p_new["Conversion"] -
      p["Conversion"]
    
    # Avoid tiny negative floating-point errors
    conversion_this_step <-
      max(conversion_this_step,
          0)
    
    # Number of additional transitions required
    additional_steps <-
      step - start_step + 1
    
    weighted_steps <-
      weighted_steps +
      additional_steps *
      conversion_this_step
    
    total_conversion_probability <-
      total_conversion_probability +
      conversion_this_step
    
    p <- p_new
  }
  
  # Conditional expectation

  if (total_conversion_probability == 0) {
    return(NA_real_)
  }
  
  weighted_steps /
    total_conversion_probability
}
```

Calculate expected steps until conversion

```{r remaining_steps}
expected_steps_all_stages <- expand_grid(
  step = 0:(max_prediction_step - 1),
  channel = channels) %>%
  
  mutate(
    expected_steps = map2_dbl(
      step,
      channel,
      ~ expected_steps_to_conversion(
        model = model_2,
        matrices = matrices_2,
        start_state = .y,
        start_step = .x,
        end_step = max_prediction_step)))
```

Plot expected steps until conversion

```{r expected_steps_plot}
expected_steps_plot_2 <- ggplot(expected_steps_all_stages,
                                aes(x = step,
                                    y = expected_steps,
                                    colour = channel)) +
  geom_line(linewidth = 1) +
  geom_vline(xintercept = max(transitions_2$step),
             linetype = "dashed") +
  labs(title = "Model 2 Expected Steps Until Conversion From Any State",
       subtitle = "Dashed Line Indicates the Maximum Observed Journey step",
       x = "Current Step",
       y = "Expected Additional Steps Until Conversion",
       colour = "Current Channel") +
  theme_minimal()

print(expected_steps_plot_2)
```

Heatmap of expected steps until conversion

```{r heatmap}
heatmap_2 <- expected_steps_all_stages %>%
  ggplot(aes(x = step,
             y = channel,
             fill = expected_steps)) +
  geom_tile() +
  scale_x_continuous(breaks = seq(0,
                                  max_prediction_step - 1,
                                  by = 10)) +
  labs(title = "Mpdel 2 Expected Steps Until Conversion From Any State",
       x = "Current Step",
       y = "Current Channel",
       fill = "Expected\nAdditional Steps") +
  theme_minimal()

print(heatmap_2)
```

Expected steps until conversion from first touchpoint

```{r steps_from_initial}
initial_2 <- expected_steps_all_stages %>%
  filter(step == 0) %>%
  arrange(expected_steps)

print(initial_2)
```

# Model Testing

Test data transitions

```{r}
#  Model 1 test transitions
transitions_1_test <- make_transitions(
  journeys_test,
  return_only = TRUE)

# Model 2 test transitions
transitions_2_test <- make_transitions(
  journeys_test,
  converting_only = TRUE)
```

Test model transition prediction accuracy

```{r test_transition_predictions}
get_test_transition_predictions <- function(
    test_data,
    model,
    prediction_max_step) {

  results <- vector("list", prediction_max_step + 1)

  # Generate a transition matrix for every step
  for (step in 0:prediction_max_step) {

    results[[step + 1]] <- make_matrix(model, step)
  }

  # Combine test observations with their predicted probabilities
  output <- test_data %>%
    filter(current_state %in% model$channels,
           next_state %in% model$states,
           step >= 0,
           step <= prediction_max_step) %>%
    rowwise() %>%
    mutate(predicted_probability =
             results[[step + 1]][current_state, next_state]) %>%
    ungroup()

  output
}
```

```{r test_predictions}
test_predictions_1 <- get_test_transition_predictions(
  transitions_1_test,
  model_1,
  max_prediction_step)

test_predictions_2 <- get_test_transition_predictions(
  transitions_2_test,
  model_2,
  max_prediction_step)
```

```{r log_loss}
calculate_log_loss <- function(predictions) {

  # Avoid log(0)
  p <- pmax(
    pmin(predictions$predicted_probability, 1 - 1e-15),
    1e-15)

  mean(-log(p))
}

log_loss_1 <- calculate_log_loss(test_predictions_1)

log_loss_2 <- calculate_log_loss(test_predictions_2)

log_loss_1

log_loss_2
```

# Conversion Rate Predictive Power

Test conversion prediction

```{r conversion_prediction_test}
# Calculate actual conversion rate for 4.1.1
actual_conversion_rate_1 <- transitions_1_test %>%
  group_by(step) %>%
  summarise(actual_conversion =
              mean(next_state == "Conversion"),
            n = n(),
    .groups = "drop")

# Model conversion rate for 4.1.1
test_initial_1 <- transitions_1_test %>%
  filter(step == 0) %>%
  count(current_state) %>%
  mutate(probability = n / sum(n))

# Propagate through mdoel 1 distributions
p <- setNames(
  rep(0, length(model_1$states)),
  model_1$states
)

p[test_initial_1$current_state] <-
  test_initial_1$probability

predicted_conversion_1 <- data.frame(
  step = 0,
  predicted_conversion = p["Conversion"])

for (step in 0:(max_prediction_step - 1)) {

  M <- make_matrix(
    model_1,
    step)

  p <- as.numeric(p %*% M)

  names(p) <- model_1$states

  predicted_conversion_1 <- bind_rows(
    predicted_conversion_1,
    data.frame(step = step + 1,
               predicted_conversion =
                 p["Conversion"]))
}
```

Test cumulative conversion prediction

```{r cumulative_conversion_test}
# Model 1 session outcomes
session_outcomes_1 <- transitions_1_test %>%
  group_by(user_id, session_id) %>%
  summarise(converted = any(next_state == "Conversion"),
            conversion_step = ifelse(
              converted,
              min(step[next_state == "Conversion"]) + 1,
              NA_real_),
            .groups = "drop")

# Actual cumulative conversion rate
actual_cumulative_1 <- tibble(step = 0:max_prediction_step) %>%
  rowwise() %>%
  mutate(actual_conversion = mean(
    session_outcomes_1$converted &
    session_outcomes_1$conversion_step <= step)) %>%
  ungroup()

# Combine with model predictions
validation_1 <- predicted_conversion_1 %>%
  left_join(actual_cumulative_1,
            by = "step") %>%
  mutate(error =
           predicted_conversion -
           actual_conversion,
         absolute_error =
           abs(error),
         squared_error =
           error^2)

# Validation metrics data range
max_test_step <- max(transitions_1_test$step,
                     na.rm = TRUE) + 1

# Validation metrics dataset
validation_metrics_1 <- validation_1 %>%
  filter(step <= max_test_step)

# Calculate validation metrics
mae_1 <- mean(abs(validation_metrics_1$predicted_conversion -
                    validation_metrics_1$actual_conversion),
              na.rm = TRUE)

rmse_1 <- sqrt(mean(
    (validation_metrics_1$predicted_conversion -
       validation_metrics_1$actual_conversion)^2,
    na.rm = TRUE))

mae_1

rmse_1
```

Plot comparison between actual and predicted conversion

```{r accuracy_plot}
# Plot
ggplot(validation_metrics_1,
       aes(x = step)) +
  geom_line(aes(y = predicted_conversion,colour = "Predicted"),
            linewidth = 1) +
  geom_point(aes(y = actual_conversion,
                 colour = "Observed")) +
  labs(title = "Model 1 Validation",
    subtitle = "Predicted vs Observed Cumulative Conversion Probability",
    x = "Current Step",
    y = "Cumulative Conversion Probability",
    colour = NULL) +
  theme_minimal()
```

# Expected Steps Predictive Power

Calculate the actual remaining steps until conversion from test data

```{r actual_remaining_steps}
# Find the actual conversion step for each converting test journey
conversion_steps_2 <- transitions_2_test %>%
  group_by(user_id, session_id) %>%
  summarise(conversion_step = min(step[next_state == "Conversion"]) + 1,
            .groups = "drop")

# Add the actual number of remaining steps to every observation in the test journeys
test_actual_2 <- transitions_2_test %>%
  left_join(conversion_steps_2,
            by = c("user_id", "session_id")) %>%
  filter(current_state %in% model_2$channels) %>%
  mutate(actual_remaining_steps =
           conversion_step - step)
```

Make test data predicted steps and true steps dataset

```{r test_step_predictions}
test_predictions_2 <- test_actual_2 %>%
  rowwise() %>%
  mutate(predicted_remaining_steps =
           expected_steps_to_conversion(
             model = model_2,
             matrices = matrices_2,
             start_state = current_state,
             start_step = step,
             end_step = max_prediction_step)) %>%
  ungroup()
```

Calculate prediction error

```{r prediction_error}
test_predictions_2 <- test_predictions_2 %>%
  mutate(error =
           predicted_remaining_steps -
           actual_remaining_steps,

    absolute_error =
      abs(error),

    squared_error =
      error^2)
```

Calculate mae and rmse

```{r validation_metrics}
mae_2 <- mean(test_predictions_2$absolute_error,
              na.rm = TRUE)

mae_2

rmse_2 <- sqrt(mean(test_predictions_2$squared_error,
                    na.rm = TRUE))

rmse_2
```

Check for prediction bias

```{r bias_check}
mean_error_2 <- mean(test_predictions_2$error,
                     na.rm = TRUE)

mean_error_2
```

Plot actual vs predicted remaining steps

```{r model_2_validation_plot}
ggplot(test_predictions_2,
       aes(x = actual_remaining_steps,
           y = predicted_remaining_steps)) +

  geom_point(alpha = 0.15) +

  geom_abline(slope = 1,
              intercept = 0,
              linetype = "dashed") +

  labs(
    title = "Model 2 Validation",
    subtitle = "Predicted vs Observed Remaining Steps to Conversion",
    x = "Actual Remaining Steps",
    y = "Predicted Remaining Steps") +

  theme_minimal()
```

Check predictions by channel

```{r validation_by_channel}
channel_validation_2 <- test_predictions_2 %>%
  group_by(current_state) %>%
  summarise(n = n(),

    mean_actual =
      mean(actual_remaining_steps),

    mean_predicted =
      mean(predicted_remaining_steps),

    MAE =
      mean(absolute_error),

    RMSE =
      sqrt(mean(squared_error)),

    mean_error =
      mean(error),

    .groups = "drop") %>%
  arrange(MAE)

channel_validation_2
```

Check predictions by step

```{r validation_by_step}
step_validation_2 <- test_predictions_2 %>%
  group_by(step) %>%
  summarise(n = n(),

    actual_mean =
      mean(actual_remaining_steps),

    predicted_mean =
      mean(predicted_remaining_steps),

    MAE =
      mean(absolute_error),

    RMSE =
      sqrt(mean(squared_error)),

    mean_error =
      mean(error),

    .groups = "drop")

step_validation_2
```

Check how well predictions are retained

```{r mae_vs_step_plot}
ggplot(step_validation_2,
       aes(x = step,
           y = MAE)) +

  geom_line() +

  geom_point() +

  labs(
    title = "Model 2 Prediction Error by Journey Step",
    x = "Current Journey Step",
    y = "Mean Absolute Error (steps)") +

  theme_minimal()
```

Plot of actual vs predicted mean steps

```{r mean_steps_plot}
step_means_2 <- test_predictions_2 %>%
  group_by(step) %>%
  summarise(actual =
              mean(actual_remaining_steps),

    predicted =
      mean(predicted_remaining_steps),

    .groups = "drop") %>%
  tidyr::pivot_longer(cols = c(actual, predicted),
                      names_to = "type",
                      values_to = "remaining_steps")

ggplot(step_means_2,
       aes(x = step,
           y = remaining_steps,
           colour = type)) +

  geom_line(linewidth = 1) +

  geom_point() +

  labs(
    title = "Model 2 Validation",
    subtitle = "Predicted vs Observed Mean Remaining Steps",
    x = "Current Journey Step",
    y = "Remaining Steps to Conversion",
    colour = NULL) +

  theme_minimal()
```

Determine the usefullness of the remaining step predictions.

```{r predictons_by_channel}
test_predictions_2 %>%
  group_by(current_state) %>%
  summarise(n = n(),
            mean_predicted = mean(predicted_remaining_steps,
                                  na.rm = TRUE),
            sd_predicted = sd(predicted_remaining_steps,
                              na.rm = TRUE),
            min_predicted = min(predicted_remaining_steps,
                                na.rm = TRUE),
            max_predicted = max(predicted_remaining_steps,
                                na.rm = TRUE),
            .groups = "drop")
```

```{r predictions_by_step}
test_predictions_2 %>%
  group_by(step) %>%
  summarise(n = n(),
            mean_predicted = mean(predicted_remaining_steps,
                                  na.rm = TRUE),
            sd_predicted = sd(predicted_remaining_steps,
                              na.rm = TRUE),
            .groups = "drop")
```

# Simulation

Model 1 simulation

```{r simulation_functions}
# Simulate one customer
simulate_customer <- function(
    matrices,
    initial_state,
    absorbing_states = c("Conversion", "Exit"),
    max_steps) {
  
  current_state <- initial_state
  
  path <- data.frame(
    step = 0,
    state = current_state)
  
  for (step in 0:(max_steps - 1)) {
    
    # Stop if already in an absorbing state
    if (current_state %in% absorbing_states) {
      break
    }
    
    # Get transition matrix for current step
    M <- matrices[[as.character(step)]]
    
    if (is.null(M)) {
      warning(
        paste("No transition matrix available for step", step))
      break
    }
    
    # Get probabilities for current state
    probabilities <- M[current_state, ]
    
    # Remove numerical problems
    probabilities[is.na(probabilities)] <- 0
    probabilities[probabilities < 0] <- 0
    
    # Normalise probabilities
    probabilities <- probabilities / sum(probabilities)
    
    # Randomly select next state
    next_state <- sample(
      names(probabilities),
      size = 1,
      prob = probabilities)
    
    # Record transition
    path <- rbind(
      path,
      data.frame(
        step = step + 1,
        state = next_state))
    
    current_state <- next_state
    
    # Stop if absorbing
    if (current_state %in% absorbing_states) {
      break
    }
  }
  
  path
}

# Simulate multiple customers
simulate_customers <- function(
    n_customers,
    matrices,
    initial_distribution,
    absorbing_states,
    max_steps) {
  
  all_paths <- vector("list", n_customers)
  
  for (customer in seq_len(n_customers)) {
    
    # Select starting state
    initial_state <- sample(
      initial_distribution$current_state,
      size = 1,
      prob = initial_distribution$probability)
    
    # Simulate journey
    path <- simulate_customer(
      matrices = matrices,
      initial_state = initial_state,
      absorbing_states = absorbing_states,
      max_steps = max_steps)
    
    path$customer_id <- customer
    
    all_paths[[customer]] <- path
  }
  
  bind_rows(all_paths)
}
```

Initial distribution based on training data

```{r simulation_1_initial}
initial_distribution_1 <- transitions_1 %>%
  filter(step == 0) %>%
  count(current_state) %>%
  mutate(probability = n / sum(n))

initial_distribution_1
```

Model 1 simulation

```{r simulation_1}
simulation_1 <- simulate_customers(
  n_customers = 10000,
  matrices = matrices_1,
  initial_distribution = initial_distribution_1,
  absorbing_states = c("Conversion", "Exit"),
  max_steps = max_prediction_step)
```

Compare simulated data to test data

```{r simulated_vs_test_conversion}
simulation_outcomes_1 <- simulation_1 %>%
  group_by(customer_id) %>%
  summarise(final_state = last(state),
            journey_length = max(step),
            .groups = "drop")

# View some simulated journey summaries
head(simulation_outcomes_1)

# Conversion rate of simulated data
mean(simulation_outcomes_1$final_state == "Conversion")

# Conversion rate from test data
test_conversion_rate <- transitions_1_test %>%
  group_by(user_id, session_id) %>%
  summarise(converted = any(next_state == "Conversion"),
            .groups = "drop") %>%
  summarise(conversion_rate = mean(converted))

test_conversion_rate
```

Model 2 simulation

```{r simulation_2_initial}
initial_distribution_2 <- transitions_2 %>%
  filter(step == 0) %>%
  count(current_state) %>%
  mutate(probability = n / sum(n))

initial_distribution_2
```

Simulation from model 2

```{r simulation_2}
simulation_2 <- simulate_customers(
  n_customers = 10000,
  matrices = matrices_2,
  initial_distribution = initial_distribution_2,
  absorbing_states = "Conversion",
  max_steps = max_prediction_step)
```

Summarise some simulated journeys

```{r simulation_2_outcomes}
simulation_outcomes_2 <- simulation_2 %>%
  group_by(customer_id) %>%
  summarise(final_state = last(state),
            journey_length = max(step),
            .groups = "drop")

# View some simulated journey summaries
head(simulation_outcomes_2)
```

Compare simulated and test data journey lengths

```{r simulated_vs_test_lengths}
# Test data journeys
test_lengths_2 <- transitions_2_test %>%
  group_by(user_id, session_id) %>%
  summarise(journey_length = max(step) + 1,
            .groups = "drop")

# Comparison table
length_comparison_2 <- data.frame(Metric = c("Number of journeys",
                                             "Mean journey length",
                                             "Median journey length",
                                             "Standard deviation",
                                             "Minimum journey length",
                                             "Maximum journey length"),
  
  Test_data = c(nrow(test_lengths_2),
                mean(test_lengths_2$journey_length),
                median(test_lengths_2$journey_length),
                sd(test_lengths_2$journey_length),
                min(test_lengths_2$journey_length),
                max(test_lengths_2$journey_length)),
  
  Simulation = c(nrow(simulation_outcomes_2),
                 mean(simulation_outcomes_2$journey_length),
                 median(simulation_outcomes_2$journey_length),
                 sd(simulation_outcomes_2$journey_length),
                 min(simulation_outcomes_2$journey_length),
                 max(simulation_outcomes_2$journey_length)))

length_comparison_2
```

Should I make a version that also uses previous journey outcome as a predictor?

Need another baseline model to compare this one to:
Have homogenous Markov model
Also use regression model of some type (maybe just GLM)