---
title: "Final Draft"
author: "Sebastian"
date: "2026-08-10"
output: html_document
---

```{r setup, include=FALSE}
setwd("D:/MXB348 - Work Integrated Learning")
customer_journey <- read.csv("D:/MXB348 - Work Integrated Learning/data/customer_journey-4.csv"); 
results <- read.csv("D:/MXB348 - Work Integrated Learning/data/results-3.csv")
### Setting the working directory and opening the files
```

```{r}
install.packages(c(
  "tidyverse",
  "markovchain"
))
### Download relevant files
```

```{r}
library(tidyverse)
library(markovchain)
### Load the relevant libraries
```

```{r}
head(customer_journey)
head(results)
str(customer_journey)
str(results)
summary(customer_journey)
summary(results)
# Displays a brief summary
```
```{r}
journeys <- customer_journey %>%
  left_join(
    results,
    by = c("user_id", "session_id")
  );
journeys <- journeys %>%
  arrange(
    user_id,
    session_id,
    order
  )
### Concatenation and rearrangement of both files to have a complete 
### understanding of a users session
### Note, the use of the sequential operator over another pipeline 
```

```{r}
View(journeys)
#  Opens viewing portal
nrow(customer_journey)
nrow(journeys)
# Debugging point, ensures equivalent dimensions
```
```{r}
transitions <- journeys %>%
  group_by(user_id, session_id) %>% # Each unique combination
  arrange(order, .by_group = TRUE) %>% # Ensures ordinal categorization
  mutate(
    next_state = lead(channel) # Looks one row ahead and puts that value into a
  ) %>% # new column called "next state"
  ungroup();# Un-magnifying
transitions <- transitions %>%
  mutate(
    next_state = case_when(
      !is.na(next_state) ~ next_state, # "na" are now converted to
      conversion == 1 ~ "Conversion",
      conversion == 0 ~ "No Conversion" 
    ) # Conversion's now act as an absorbing state in a Markov Chain.
  ) # To symbolize the end of an individual tracked session.
View(transitions %>%
    select(
      user_id,
      session_id,
      order,
      channel,
      next_state,
      conversion
    )
)
### Formalizing the Markov-chain transformation mechanism (Markov Property), 
### by identifying what channel each customer moves to next.
```

```{r}
transition_counts <- transitions %>%
  count(
    channel,
    next_state,
    name = "n"
  );
View(transition_counts)
### This is the point in which we can begin to really interpret our data. Since
### now we have sample count data associated with each possible unique tracked 
### transition for every person and each of their sessions. 
```

```{r}
transition_probs <- transition_counts %>%
  group_by(channel) %>%
  mutate(
    probability = n / sum(n)
  ) %>%
  ungroup()
View(transition_probs)
### Converts the count data into probabilities
```

```{r}
transition_probs %>%
  filter(next_state == "Conversion") %>%
  arrange(desc(probability))
### A tibble that shows the probability given the sample data of each channel
### turning into a conversion.
```
```{r}
states <- sort(
  unique(
    c(
      transitions$channel,
      transitions$next_state
    )
  )
);
P <- matrix(
  0,
  nrow = length(states),
  ncol = length(states),
  dimnames = list(states, states) # a R^(10x10) matrix with 
) # [channel no conversion conversion].
### This is all to create an empty/unfilled transition probability matrix where 
### after this section we will fill it with data that satisfy the necessary 
### properties (Square matrix with probabilities that sum to 1).
```

```{r}
for (i in seq_len(nrow(transition_probs))) {

  from <- transition_probs$channel[i]
  to   <- transition_probs$next_state[i]

  P[from, to] <-
    transition_probs$probability[i]
};
P["Conversion", ] <- 0
P["Conversion", "Conversion"] <- 1

P["No Conversion", ] <- 0
P["No Conversion", "No Conversion"] <- 1
```

```{r}
customer_mc <- new(
  "markovchain",
  states = states,
  transitionMatrix = P,
  name = "Logickube Customer Journey"
);
customer_mc;
View(P)
### We have now constructed our first-order Markov Chain transition probability
### matrix. We have also got a probability of 1 in both the "No Conversion" and
### "Conversion" diagonal to represent them as absorbing states that symbolize
### how at the end of every session, we end up with either a successful or 
### unsuccessful conversion.
### Instagram has the highest observed immediate transition probability 
### conversion at approximately 25.64% with YouTube being the lowest at 25.09%.
```

### Session Identification variable (user_id) (Ordinal Discrete Integers)
### A greater number simulates a later discrete time period
### Session vs Order
### A single session is a single period of tracked continuity. In essence, discrete differences in sessions within the same user shows a discontinuity. Where the user has spent x amount of time (x could be seconds, minutes, hours or days since the last session)

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
