---
title: "Final Draft"
author: "Sebastian"
date: "2026-08-10"
output: html_document
---

```{r setup, include=FALSE}
setwd("D:/MXB348 - Work Integrated Learning")
customer_journey <- read.csv("D:/MXB348 - Work Integrated Learning/data/customer_journey-4.csv"); 
results <- read.csv("D:/MXB348 - Work Integrated Learning/data/results-3.csv")
### Setting the working directory and opening the files
```

```{r}
install.packages(c(
  "tidyverse",
  "markovchain"
))
### Download relevant files
```


```{r}
library(tidyverse)
library(markovchain)
### Load the relevant libraries
```


```{r}
head(customer_journey)
head(results)
str(customer_journey)
str(results)
summary(customer_journey)
summary(results)
# Disp
```
```{r}
journeys <- customer_journey %>%
  left_join(
    results,
    by = c("user_id", "session_id")
  )
```


```{r}
View(journeys)
```

head(customer_journey)
### User Identification variable (user_id) (Discrete Integers)
unique_id <- unique(customer_journey$user_id)
Unique_Users = length(unique_id)
### Session Identification variable (user_id) (Ordinal Discrete Integers)
### A greater number simulates a later discrete time period
### Session vs Order
### A single session is a single period of tracked continuity. In essence, discrete differences in sessions within the same user shows a discontinuity. Where the user has spent x amount of time (x could be seconds, minutes, hours or days since the last session)

```

```{r}
str(customer_journey)
str(results)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
