---
title: "Final Draft"
author: "Sebastian"
date: "2026-08-10"
output: html_document
---

```{r setup, include=FALSE}
setwd("D:/MXB348 - Work Integrated Learning")
customer_journey <- read.csv("D:/MXB348 - Work Integrated Learning/data/customer_journey-4.csv"); read.csv("D:/MXB348 - Work Integrated Learning/data/results-3.csv")

```


head(customer_journey)
### User Identification variable (user_id) (Discrete Integers)
unique_id <- unique(customer_journey$user_id)
Unique_Users = length(unique_id)
### Session Identification variable (user_id) (Ordinal Discrete Integers)
### A greater number simulates a later discrete time period
### Session vs Order
### A single session is a single period of tracked continuity. In essence, discrete differences in sessions within the same user shows a discontinuity. Where the user has spent x amount of time (x could be seconds, minutes, hours or days since the last session)

```

```{r}
results <- read.csv("D:\Semester 2 2026\MXB348 - Work Integrated Learning/data/customer_journey-4.csv)
head(results)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
