---
title: "Final Draft"
author: "Sebastian"
date: "2026-08-10"
output: html_document
---

```{r setup, include=FALSE}
setwd("D:/MXB348 - Work Integrated Learning")
customer_journey <- read.csv("D:/MXB348 - Work Integrated Learning/data/customer_journey-4.csv"); 
results <- read.csv("D:/MXB348 - Work Integrated Learning/data/results-3.csv")
### Setting the working directory and opening the files
```

```{r}
install.packages(c(
  "tidyverse",
  "markovchain"
))
### Download relevant files
```

```{r}
library(tidyverse)
library(markovchain)
### Load the relevant libraries
```

```{r}
head(customer_journey)
head(results)
str(customer_journey)
str(results)
summary(customer_journey)
summary(results)
# Displays a brief summary
```
```{r}
journeys <- customer_journey %>%
  left_join(
    results,
    by = c("user_id", "session_id")
  );
journeys <- journeys %>%
  arrange(
    user_id,
    session_id,
    order
  )
### Concatenation and rearrangement of both files to have a complete 
### understanding of a users session
### Note, the use of the sequential operator over another pipeline 
```

```{r}
View(journeys)
#  Opens viewing portal
nrow(customer_journey)
nrow(journeys)
# Debugging point, ensures equivalent dimensions
```
```{r}
transitions <- journeys %>%
  group_by(user_id, session_id) %>% # Each unique combination
  arrange(order, .by_group = TRUE) %>% # Ensures ordinal categorization
  mutate(
    next_state = lead(channel) # Looks one row ahead and puts that value into a
  ) %>% # new column called "next state"
  ungroup() # Un-magnifying
### Formalizing the Markov-chain transformation mechanism (Markov Property), 
### by identifying what channel each customer moves to next.
```

```{r}

```


### Session Identification variable (user_id) (Ordinal Discrete Integers)
### A greater number simulates a later discrete time period
### Session vs Order
### A single session is a single period of tracked continuity. In essence, discrete differences in sessions within the same user shows a discontinuity. Where the user has spent x amount of time (x could be seconds, minutes, hours or days since the last session)

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
