---
title: "Question_3"
author: "Retief Padget"
date: "2026-09-03"
output: pdf_document
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=F}
library(ChannelAttribution)
library(tidyverse)
  # More Specifically
  library(dplyr)
  library(lubridate)
  library(stringr)
library(markovchain)
library(data.table)
```

```{r data_load, include=F}
journey <- read_csv("C:/Uni26_Sem2/MXB348/WIL/MXB348/data/customer_journey-4.csv")
results <- read_csv("C:/Uni26_Sem2/MXB348/WIL/MXB348/data/results-3.csv")

## FEATURE ENGINEERING
  # path with same order id into 1 column.
  # can look at frequency of each path.
  # group into journey length too.
  # duration time + grouped too.
  # compare with collapsed group.
  # looking at total conversion rate and then provide weighting to EACH CHANNEL.

# direction
  # no grouping
  # grouping - time, channels, collapsing.
```

## 4.3 [OUTLINE] Business Success Attribution

Business Success = Conversion. Attribution method determines how much credit each contributing factor deserves.

### 4.3.1 Contribution of Each Channel

Calculate OVERALL conversion rate & Attribute the conversions to the 8 channels.

Example: calculate 8 weights (1 for each channel) that sum up to 100%.

### 4.3.2 Journey Length Impact

Does the customer journey length have an impact on attribution results?

Ideas: Short vs long. Duration of time.

Do the attribution results differ for collapsed journeys?

### 4.3.3 Levels of Granularity

8 channels to 3 groups: Free (organic, direct, email); Google (search, display, youtube); Meta (facebook, instagram)

Are attribution results CONSISTENT between the two levels of granularity?

Do attribution results at Level 2 (no group) add up to results at Level 1 (groups)? If not, is there any method?

## [LITERATURE REVIEW]

## [CLEANING / FEATURE ENGINEERING]

```{r feature engineering}
# CHANNEL NAME SYSTEM
journeyA <- journey %>%
  mutate(channel = fct_recode(channel,
                             "GD" = "Google Display",
                             "GS" = "Google Search",
                             "FO" = "Organic",
                             "FD" = "Direct",
                             "GY" = "Youtube",
                             "MI" = "Instagram",
                             "MF" = "Facebook",
                             "FE" = "Email Marketing")) # Codename: 1st Letter = Group, 2nd Letter = Channel

# CHANNEL JOURNEY (WHOLE)
journeyA <- setDT(journeyA)
journeyA_wide = dcast(data = journeyA, formula = user_id + session_id ~ order, value.var = "channel")
journeyA_wide = journeyA_wide[, path:=do.call(paste,c(.SD, sep=" > ")), .SDcols=-c(1,2)]
journeyA_wide = journeyA_wide[, path:=word(path, 1, sep = " > NA")]
journeyA_wide <- journeyA_wide %>% select(-c("0":"25"))

# FIRST + LAST CHANNEL
journeyA_wide <- journeyA_wide %>% 
  mutate("first_touchpoint" = str_extract(path, "^.."),
         "last_touchpoint" = str_extract(path, "..$"))

# TOTAL JOURNEY TIME
journeyA <- journeyA %>%
  filter(order == 0) %>%
  select(-c("order", "channel"))

merge1 <- inner_join(journeyA_wide, journeyA, by = c("user_id", "session_id"))
attribution_v1 <- inner_join(merge1, results, by = c("user_id", "session_id"))

# TOTAL TOUCHPOINT COUNT
n_touchpoint <- attribution_v1$path %>%
  str_count(">") + 1

# Make a new column in our dataset
attribution_v1$touchpoint_count <- n_touchpoint

# TOTAL CONVERSIONS (from first to current customer journey)
attribution_v1 <- attribution_v1 %>%
  group_by(user_id) %>%
  mutate(customer_conversion_count = cumsum(conversion == 1),
         customer_journey_count = session_id + 1,
         customer_conversion_rate = round(customer_conversion_count/customer_journey_count, 3),
         returning = n() > 1) %>%
  ungroup()

```
