---
title: "Question_3"
author: "Retief Padget"
date: "2026-09-03"
output: pdf_document
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=F}
library(ChannelAttribution)
library(tidyverse)
  # More Specifically
  library(dplyr)
  library(lubridate)
  library(stringr)
library(markovchain)
library(data.table)
library(reshape2)
library(igraph)
```

```{r data_load, include=F}
journey <- read_csv("data/customer_journey-4.csv")
results <- read_csv("data/results-3.csv")

## FEATURE ENGINEERING
  # path with same order id into 1 column.
  # can look at frequency of each path.
  # group into journey length too.
  # duration time + grouped too.
  # compare with collapsed group.
  # looking at total conversion rate and then provide weighting to EACH CHANNEL.

# direction
  # no grouping
  # grouping - time, channels, collapsing.
```

## 4.3 [OUTLINE] Business Success Attribution

Business Success = Conversion. Attribution method determines how much credit each contributing factor deserves.

### 4.3.1 Contribution of Each Channel

Calculate OVERALL conversion rate & Attribute the conversions to the 8 channels.

Example: calculate 8 weights (1 for each channel) that sum up to 100%.

### 4.3.2 Journey Length Impact

Does the customer journey length have an impact on attribution results?

Ideas: Short vs long. Duration of time.

Do the attribution results differ for collapsed journeys?

### 4.3.3 Levels of Granularity

8 channels to 3 groups: Free (organic, direct, email); Google (search, display, youtube); Meta (facebook, instagram)

Are attribution results CONSISTENT between the two levels of granularity?

Do attribution results at Level 2 (no group) add up to results at Level 1 (groups)? If not, is there any method?

## [LITERATURE REVIEW]

## [CLEANING / FEATURE ENGINEERING]

```{r feature engineering}
# CHANNEL NAME SYSTEM
journeyA <- journey %>%
  mutate(channel = fct_recode(channel,
                             "GD" = "Google Display",
                             "GS" = "Google Search",
                             "FO" = "Organic",
                             "FD" = "Direct",
                             "GY" = "Youtube",
                             "MI" = "Instagram",
                             "MF" = "Facebook",
                             "FE" = "Email Marketing")) # Codename: 1st Letter = Group, 2nd Letter = Channel

# CHANNEL JOURNEY (WHOLE)
journeyA <- setDT(journeyA)
journeyA_wide = data.table::dcast(data = journeyA, formula = user_id + session_id ~ order, value.var = "channel")
  # Clarify that dcast is from "data.table". Assumes "reshape2" if both installed.
journeyA_wide = journeyA_wide[, path:=do.call(paste,c(.SD, sep=" > ")), .SDcols=-c(1,2)]
journeyA_wide = journeyA_wide[, path:=word(path, 1, sep = " > NA")]
journeyA_wide <- journeyA_wide %>% select(-c("0":"25"))

# FIRST + LAST CHANNEL
journeyA_wide <- journeyA_wide %>% 
  mutate("first_touchpoint" = str_extract(path, "^.."),
         "last_touchpoint" = str_extract(path, "..$"))

# TOTAL JOURNEY TIME
journeyA <- journeyA %>%
  filter(order == 0) %>%
  select(-c("order", "channel"))

merge1 <- inner_join(journeyA_wide, journeyA, by = c("user_id", "session_id"))
attribution_v1 <- inner_join(merge1, results, by = c("user_id", "session_id"))

# TOTAL TOUCHPOINT COUNT
n_touchpoint <- attribution_v1$path %>%
  str_count(">") + 1

# Make a new column in our dataset
attribution_v1$touchpoint_count <- n_touchpoint

# TOTAL CONVERSIONS (from first to current customer journey)
attribution_v1 <- attribution_v1 %>%
  group_by(user_id) %>%
  mutate(customer_conversion_count = cumsum(conversion == 1),
         customer_journey_count = session_id + 1,
         customer_conversion_rate = round(customer_conversion_count/customer_journey_count, 3),
         returning = n() > 1) %>%
  ungroup()

# NO CONVERSION (For Markov Model)
attribution_v2 <- attribution_v1 %>%
  mutate(null_conversion = 1 - conversion)

```

```{r attribution model}
# calculating the models
mod1 <- markov_model(attribution_v2,
                     var_path = 'path',
                     var_conv = 'conversion',
                     var_null = 'null_conversion',
                     out_more = TRUE)
# extracting the results of attribution
df_res1 <- mod1$result
# extracting a transition matrix
df_trans1 <- mod1$transition_matrix
# naming
df_trans1 <- df_trans1 %>%
  mutate(channel_from = fct_recode(channel_from,
                                   "GD" = "1",
                                   "GS" = "2",
                                   "FO" = "3",
                                   "FD" = "4",
                                   "GY" = "5",
                                   "MI" = "6",
                                   "MF" = "7",
                                   "FE" = "8"),
         channel_to = fct_recode(channel_to,
                                 "GD" = "1",
                                 "GS" = "2",
                                 "FO" = "3",
                                 "FD" = "4",
                                 "GY" = "5",
                                 "MI" = "6",
                                 "MF" = "7",
                                 "FE" = "8"))
df_trans1 <- dcast(df_trans1, channel_from ~ channel_to, value.var = 'transition_probability')

### plotting the Markov graph ###
df_trans <- mod1$transition_matrix
# adding dummies in order to plot the graph
df_dummy <- data.frame(channel_from = c('(start)', '(conversion)', '(null)'),
                       channel_to = c('(start)', '(conversion)', '(null)'),
                       transition_probability = c(0, 1, 1))
df_trans <- rbind(df_trans, df_dummy)
# name editing
df_trans <- df_trans %>%
  mutate(channel_from = fct_recode(channel_from,
                                   "GD" = "1",
                                   "GS" = "2",
                                   "FO" = "3",
                                   "FD" = "4",
                                   "GY" = "5",
                                   "MI" = "6",
                                   "MF" = "7",
                                   "FE" = "8"),
         channel_to = fct_recode(channel_to,
                                 "GD" = "1",
                                 "GS" = "2",
                                 "FO" = "3",
                                 "FD" = "4",
                                 "GY" = "5",
                                 "MI" = "6",
                                 "MF" = "7",
                                 "FE" = "8"))
# ordering channels
df_trans$channel_from <- factor(df_trans$channel_from,
                                levels = c('(start)', '(conversion)', '(null)', 'GD', 'GS', 'FO', 'FD', 'GY', 'MI', 'MF', 'FE'))
df_trans$channel_to <- factor(df_trans$channel_to,
                                levels = c('(start)', '(conversion)', '(null)', 'GD', 'GS', 'FO', 'FD', 'GY', 'MI', 'MF', 'FE'))
df_trans <- dcast(df_trans, channel_from ~ channel_to, value.var = 'transition_probability')

# creating the markovchain object
trans_matrix <- matrix(data = as.matrix(df_trans[, -1]),
nrow = nrow(df_trans[, -1]), ncol = ncol(df_trans[, -1]),
dimnames = list(c(as.character(df_trans[, 1])), c(colnames(df_trans[, -1]))))
trans_matrix[is.na(trans_matrix)] <- 0
trans_matrix1 <- new("markovchain", transitionMatrix = trans_matrix)
# plotting the graph
plot(trans_matrix1, edge.arrow.size = 0.35, margin = -0.3)

g <- as(trans_matrix1, "igraph") 
lo <- layout_with_fr(g) # Fruchterman-Reingold layout
lo_scaled <- lo * 1.5   # Increase distance by 50%
plot(g, layout = lo_scaled, edge.arrow.size = 0.35)
```
