---
title: "Question_3"
author: "Retief Padget"
date: "2026-09-03"
output: pdf_document
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=F}
library(ChannelAttribution)
library(tidyverse)
  # More Specifically
  library(dplyr)
  library(lubridate)
  library(stringr)
library(markovchain)
```

```{r data_load, include=F}
journey <- read_csv("data/customer_journey-4.csv")
results <- read_csv("data/results-3.csv")

## FEATURE ENGINEERING
  # path with same order id into 1 column.
  # can look at frequency of each path.
  # group into journey length too.
  # duration time + grouped too.
  # compare with collapsed group.
  # looking at total conversion rate and then provide weighting to EACH CHANNEL.

# direction
  # no grouping
  # grouping - time, channels, collapsing.
```

## 4.3 [OUTLINE] Business Success Attribution

Business Success = Conversion. Attribution method determines how much credit each contributing factor deserves.

### 4.3.1 Contribution of Each Channel

Calculate OVERALL conversion rate & Attribute the conversions to the 8 channels.

Example: calculate 8 weights (1 for each channel) that sum up to 100%.

### 4.3.2 Journey Length Impact

Does the customer journey length have an impact on attribution results?

Ideas: Short vs long. Duration of time.

Do the attribution results differ for collapsed journeys?

### 4.3.3 Levels of Granularity

8 channels to 3 groups: Free (organic, direct, email); Google (search, display, youtube); Meta (facebook, instagram)

Are attribution results CONSISTENT between the two levels of granularity?

Do attribution results at Level 2 (no group) add up to results at Level 1 (groups)? If not, is there any method?

## [LITERATURE REVIEW]

## [CLEANING / FEATURE ENGINEERING]

```{r path journey column}
journey2 <- journey %>%
  mutate(channel = fct_recode(channel,
                             "GD" = "Google Display",
                             "GS" = "Google Search",
                             "FO" = "Organic",
                             "FD" = "Direct",
                             "GY" = "Youtube",
                             "MI" = "Instagram",
                             "MF" = "Facebook",
                             "FE" = "Email Marketing")) %>% # Codename: 1st Letter = Group, 2nd Letter = Channel

dt_wide = dcast(data = journey2, formula = user_id ~ user_id, value.var = "channel")
dt_wide = dt_wide[, path:=do.call(paste,c(.SD, sep=' > ')), .SDcols=-1]
dt_wide = dt_wide[, path:=word(path, 1, sep = " > NA")]

conversion = dt[, .(conversion=sum(conversion), conversion_value=sum(conversion_value)), by=cookie]

setkey(conversion, cookie)
setkey(dt_wide, cookie)

dt_wide = merge(dt_wide, conversion)

head(dt_wide[, .(path, conversion, conversion_value)]) %>%
           kable() %>%
           kable_styling(bootstrap_options = "striped",
                         full_width = F, 
                         position="center")

```
